---
title: "Playing Simple Games with Reinforcement Learning"
author:
- name: Sean Xiaohao Duan
  affiliation: '1'
  email: sxdff5@mail.missouri.edu
- name: David Reynolds
  affiliation: '1'
  email: davidreynolds@mail.missouri.edu
subtitle: A Final Project for Data IV
output: pdf_document
affiliation:
- id: '1'
  institution: University of Missouri
---

```{r setup}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, include = FALSE)
knitr::knit_hooks$set(purl = knitr::hook_purl)
```


```{r pkgload, include = FALSE}
#PUT YOUR LIBRARY PACKAGES HERE WHEN YOU ARE READY

#sean's pkgs
library(lme4)
library(ggplot2)
library(papaja)
library(sjPlot)
library(tidyverse)
library(nnet)
```

## Introduction

For our final project for Data IV, our group decided to use the Reinforcement Learning Methods that we had learned previously in this semester of class. We believed that this was a good choice, as it seemed direct and simple to implement, with a large variety of potential applications. We decided as a group that some simple traditional games would be a good field of application for our reinforcement learning methods. We explain further our purposes and choices of methodology in the following sections focusing on each of our individual efforts in this group project.

```{r child = "david_project.Rmd"}
```


```{r child = "sean_project.Rmd"}
```


## Conclusion

