---
title: '**Playing simple games with reinforcement learning**'
author:
- name: '*Sean Duan and David Reynolds, University of Missouri*'
output:
  html_document:
    df_print: paged
---

```{r pkgload, include = FALSE}
# Sean's pkgs
library(lme4)
library(ggplot2)
library(papaja)
library(sjPlot)
library(tidyverse)
library(nnet)
```

For our final project for STAT 9340: Data Analysis IV, we decided to focus on reinforcement learning. We believe this was a good choice, as reinforcement learning is relatively intuitive and has a variety of applications. We decided that some simple traditional games would be a good area of application, specifically rock paper scissors and Connect Four. We explain further our motivation and methodology in the following sections, each focusing on one of the two games.

```{r child = "sean_project.Rmd"}
```

```{r child = "david_project.Rmd"}
```

### **Overall project conclusion** 

This project was an extremely satisfying opportunity to grow as students and statisticians. It is always important to try and see if we can grow and adapt to new problems, and getting to choose our own area of focus for this last project led to ample learning. One of the biggest problems we dealt with was efficiently collaborating across two different programming languages on two separate but related problems. We used GitHub for collaboration, and as compared to our previous project in Data III (which our group members also worked on together), we had a significantly easier experience, reflective of our increased mastery of the platform. Additionally, integrating images of Python output and our R figures was a challenge but a worthwhile one with regard to producing a final report and product that was relatively easy to use.
